---
title: "Practicing Regular Expressions"
subtitle: "Math 241, Week 8"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
#Load tidyverse
library(tidyverse)
library(mdsr)
```


## Problem 1 (Easy): 

Use the `Macbeth_raw` data from the `mdsr` package to answer the following questions:

a. Find all of the adjectives in Macbeth that end in more or less


```{r}
data(Macbeth_raw)

macbeth <- Macbeth_raw %>%
  str_split(" ") %>%
  pluck(1)

macbethMoreLess <- macbeth %>%
  str_subset("[a-z](more|less)$")
```


b. Find all of the lines containing the stage direction Exit or Exeunt.

```{r}
data(Macbeth_raw)

macbeth2 <- Macbeth_raw %>%
  str_split("\r\n") %>%
  pluck(1)

macbethExit <- macbeth2 %>%
  str_subset("(Exit|Exeunt)")
```

## Problem 2 (Medium): 

a. Use regular expressions to determine the number of speaking lines in the Complete Works of William Shakespeare. Here, we care only about how many times a character speaks—not what they say or for how long they
speak.

```{r, eval = FALSE}
shakespeare_url <- "https://www.gutenberg.org/files/100/100-0.txt"
shakespeare <- RCurl::getURL(shakespeare_url)
view(shakespeare)
```

```{r}
shakespeare_chars <- tribble(~name, ~regexp, "")
```

b. Make a bar chart displaying the top 100 characters with the greatest number of lines. Hint you may want to use either the `stringr::str_extract` or `strsplit` function here.

```{r}
# Here are two different ways to get the list of speakers:

```


